---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# scATACutils: an R/Bioconductor package for working with 10x scATACseq data


<!-- badges: start -->
<!-- badges: end -->

The goal of scATACutils is to provide functions to work with 10x scATACseq data. It includes functions to calculate qualtiy control metrics such as banding score, Frip score,  and TSS enrichment score etc. Some convinient functions are provided for visulizations as well. Plot the scatter plot of the Frip and read depth. Plot scATACseq coverage tracks by group etc.

## Installation

You can install the released version of scATACutils from github with:

``` r
devtools::install_github("crazyhottommy/scATACutils")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(scATACutils)
## basic example code
```

